---
title: "stats10_assignment2_p1"
output: html_document
date: "2023-04-25"
---

Name: Vedang Bhelande

================== PART 1 ==================

Exercise 1

a) Download the data from the course site and read it into R. When you read in the data, name your object “flint”.

```{r}
flint = read.csv("/Users/vedangbhelande/Desktop/UCLA SQ '23/Stats 10/Assignments/Assignment 2/flint.csv")
```

b) The EPA states a water source is especially dangerous if the lead level is 15 PPB or greater. What proportion of the locations tested were found to have dangerous lead levels?

```{r}
# Read in column containing lead levels
leadLevels = flint$Pb

# Print statement to check we are reading in Pb levels
# print(leadLevels)

# Proportion of locations where lead levels were 15 PPB or greater
dangSources = mean(leadLevels >= 15)
print(dangSources)
```

c) Report the mean copper level for only test sites in the North region.

```{r}
# Read in column containing copper levels
copperLevels = flint$Cu

# Read in column containing region
region = flint$Region

# Subset copper levels based on region
meanCuNorth = mean(copperLevels[region == "North"])
paste("Mean copper level for only test sites in the North region was", meanCuNorth, "PPB.")
```

d) Report the mean copper level for only test sites with dangerous lead levels (at least 15 PPB).

```{r}
# Subset copper levels based on lead levels
meanCu_dangPb = mean(copperLevels[leadLevels >= 15])
paste("Mean copper level for only test sites with dangerous lead levels (at least 15 PPB) was", meanCu_dangPb, "PPB.")
```

e) Report the mean lead and copper levels.

```{r}
meanPbLevel = mean(leadLevels)
paste("Mean lead level is", meanPbLevel, "PPB.")

meanCuLevel = mean(copperLevels)
paste("Mean copper level is", meanCuLevel, "PPB.")
```

f) Create a box plot with a good title for the lead levels.

```{r}
# install.packages("mosaic")
library(mosaic)

# safePbLevel = (leadLevels < 15)
# dangPbLevel = (leadLevels >= 15)

# Create box plot woth nice title
boxplot(leadLevels, main = "Lead Levels in Flint (PPB)")

```

g) Based on what you see in part (f), does the mean seem to be a good measure of center for the data? Report a more useful statistic for this data.

```{r}
# View 5 number summary
summary(leadLevels)

mean(leadLevels == 0)
median(leadLevels)
```

The mean is not a good measure of center since the box plot indicates a positively skewed distribution and there are a large number of outliers. In this case, the median is a more useful statistic and has a value 0.

=============================

Exercise 2

a) Construct a scatterplot of Life against Income. Note: Income should be on the horizontal axis. How does income appear to affect life expectancy?

```{r}
# Read in data
life<-read.table("/Users/vedangbhelande/Desktop/UCLA SQ '23/Stats 10/Assignments/Assignment 2/countries_life.txt", header = TRUE)

lifeExpectancy = life$Life
income = life$Income

plot(income, lifeExpectancy, xlab = "Income ($)", ylab = "Life Expectancy (Years)", main = "Life Expectancy v/s Income", col = "red")
```

Higher income seems to be associated with higher life expectancy. After a while, ...

b) Construct the boxplot and histogram of Income. Are there any outliers?

```{r}
#Boxplot of Income
boxplot(income)
```

```{r}
# Histogram of Income
hist(income)
```

From the boxplot, anything after \$2500 seems to be an outlier. From histogram, anything after \$1000 seems to be an outlier.

c) Split the data set into two parts: One for which the Income is strictly below \$1000, and one for which the Income is at least \$1000. Come up with your own names for these two objects.

```{r}
# Income strictly below $1000
incomeBelow1k = income[income < 1000]
lifeBelow1k = lifeExpectancy[income < 1000]
# print(incomeBelow1k)

# Income is at least $1000
incomeAbove1k = income[income >= 1000]
lifeAbove1k = lifeExpectancy[income >= 1000]
# print(incomeAbove1k)
```

d) Use the data for which the Income is below $1000. Plot Life against Income and compute the correlation coefficient. Hint: use the function cor().

```{r}
plot(incomeBelow1k, lifeBelow1k, xlab = "Income ($)", ylab = "Life Expectancy (Years)", main = "Life Expectancy v/s Income", col = "red")
```

=============================

Exercise 3

a) Compute the summary statistics for lead and zinc using the summary() function.

```{r}
# Read in data
maas<-read.table("/Users/vedangbhelande/Desktop/UCLA SQ '23/Stats 10/Assignments/Assignment 2/soil.txt", header = TRUE)

# Lead summary
leadConc = maas$lead
print("Summary statistics for lead is as follows:")
summary(leadConc)

# Zinc Summary
zincConc = maas$zinc
print("Summary statistics for zinc is as follows:")
summary(zincConc)
```

Plot two histograms: one of lead and one of log(lead).

```{r}
# Histogram of lead
hist(leadConc)
```

```{r}
# Histogram of log(lead)
logPb = log(leadConc)
hist(logPb)
```

c) Plot log(lead) against log(zinc). What do you observe?

```{r}
# log(zinc) Conversion
logZn = log(zincConc)

# Scatterplot of log(lead) against log(zinc)
plot(logZn, logPb, xlab = "log(zinc)", ylab = "log(lead)", main = "log(lead) v/s log(zinc)")
```

d) The level of risk for surface soil based on lead concentration in ppm is given on the table below:

Mean concentration (ppm)        Level of risk
Below 150                       Lead-free
Between 150-400                 Lead-safe
Above 400                       Significant environmental lead hazard

Use techniques similar to last lab to give different colors and sizes to the lead concentration at these 155 locations. You do not need to use the maps package create a map of the area. Just plot the points without a map.

```{r}
# Make cuts
PbCuts = cut(leadConc, c(0, 150, 400, 700))

# PbFree = leadConc[leadConc < 150]
# PbSafe = leadConc[leadConc >= 150 & leadConc < 400]
# print(PbSafe)
# PbDang = leadConc[leadConc >= 400]

# Assign colors
myColors = c("red", "orange", "green")
PbColors = myColors[as.numeric(PbCuts)]

# Assign sizes
mySizes = c(1, 4, 9)

# Plot
plot(x = maas$x, y = maas$y, xlab = " x Coordinate", ylab = "y Coordinate", main = " Lead Levels at Different Coordinates", col = PbColors, pch = 19, cex = leadConc/mean(leadConc))
```

=============================

Exercise 4

a) Plot the data point locations. Use good formatting for the axes and title. Then add the outline of LA County by typing:

```{r}
# Install maps package
# install.packages("maps")
library(maps)

# Read in data
LA<-read.table("/Users/vedangbhelande/Desktop/UCLA SQ '23/Stats 10/Assignments/Assignment 2/la_data.txt", header = TRUE)

# Plot data locations
lat = LA$Latitude
long = LA$Longitude
plot(long, lat, xlim = c(-119, -118), ylim = c(33.5, 34.5))

# Adding outline
map("county", "california", add = TRUE)
```

b) Do you see any relationship between income and school performance? Hint: Plot the variable Schools against the variable Income and describe what you see. Ignore the data points on the plot for which Schools = 0. Use what you learned about subsetting with logical statements to first create the objects you need for the scatter plot. Then, create the scatter plot.

```{r}
# Remove values where Schools = 0
LAnon0 = LA[LA$Schools != 0, ]

plot(LAnon0$Income, LAnon0$Schools)
```

We see a positive relation ship between income and schools until around $100,000. After that the relationship vanishes and we see outliers which indicates income stops mattering after that point.

=============================

Exercise 5

```{r}
# Read in data
customer_data<-read.csv("/Users/vedangbhelande/Desktop/UCLA SQ '23/Stats 10/Assignments/Assignment 2/customer_data.csv", header = TRUE)
```

a) Are there any missing values in the dataset? If so, how many are there and which variables have missing values?

```{r}
# Remove missing values
na.omit(customer_data)
```

```{r}
# Count number of missing values
paste("The total number of missing values is", sum(is.na(customer_data)))

# Tell which categories have missing values
paste("The number of missing values in customer ID is", sum(is.na(customer_data$cust_id)))
paste("The number of missing values in age is", sum(is.na(customer_data$age)))
paste("The number of missing values in gender is", sum(is.na(customer_data$gender)))
paste("The number of missing values in income is", sum(is.na(customer_data$income)))
paste("The number of missing values in education is", sum(is.na(customer_data$education)))
paste("The number of missing values in marital status is", sum(is.na(customer_data$marital_status)))
paste("The number of missing values in purchase amount is", sum(is.na(customer_data$purchase_amt)))
```

b) What is the data type of each variable? Are there any variables that should be converted to a different data type?

```{r}
# Head shows data type of each variable
head(customer_data)
```

Customer ID is character.
Age is integer.
Gender is character.
Income is integer.
Education is character.
Marital Status is character.
Purchase Amount is integer.

```{r}
# Factor gender to convert characters into items with unique integer representations
factor(customer_data$gender)

# Factor marital status
factor(customer_data$marital_status)
```

We can encode the categorical variables of gender and marital status numerically for easier data analysis in the future.

c) Do any numerical variables have outliers or extreme values? If so, how would you handle them? Provide your analysis in R for identifying outliers (e.g., visualization, numerical summary statistics). This is an open-ended question, so please feel free to use any appropriate methods to identify and deal with any outliers or extreme values in the dataset.

The approach used here is to make boxplots and check the 5 number summary for each numeric variable and check for any outliers.

```{r}
# Age Boxplot
boxplot(customer_data$age)
```
```{r}
# Age 5 number summary
summary(customer_data$age)
max_age = 56.75 + 1.5*(56.75-32.00)
min_age = 32.00 - 1.5*(56.75-32.00)
paste("Maximum age that is not an outlier <=", max_age)
paste("Minimum age that is not an outlier >=", min_age)
```

From the boxplot and 5 number summary, we see no age values that can be an outlier.

```{r}
# Purchase Amount Boxplot
boxplot(customer_data$purchase_amt)
```
```{r}
# Purchase Amount 5 Number Summary
summary(customer_data$purchase_amt)
max_purchase = 466.0 + 1.5*(466.0-211.0)
min_purchase = 211.0 - 1.5*(466.0-211.0)
paste("Maximum purchase amount that is not an outlier <=", max_purchase)
paste("Minimum purchase amount that is not an outlier >=", min_purchase)
```

From the boxplot and 5 number summary, we see no purchase amount values that can be an outlier.

```{r}
# Income Boxplot
boxplot(customer_data$income)
```

```{r}
# Purchase Amount 5 Number Summary
summary(customer_data$income)
max_inc = 150030 + 1.5*(150030-55320)
min_inc = 55320 - 1.5*(150030-55320)
paste("Maximum income that is not an outlier <=", max_inc)
paste("Minimum income that is not an outlier >=", min_inc)
```

From the boxplot and 5 number summary, we see no income values that can be an outlier.